Spin canonical divisors of log stacky curves

### 2015-06-24 stacky_write_up/spin_cring.tex
some of proof g=0 main thm, minor edits

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has relations in degree at most $\max(11, e_1,
\ldots, e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has relations in degree at most
$\max(10, e_1, \ldots, e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g-0-exceptional}
for signatures with $g = 0$.
\end{thm}
\todo{Say that this follows immediately from the three theorems for the special cases.}
Note that in the course of the proof of Theorem ~\ref{thm:g_0_generators_relations} we will not merely obtain bounds on the degrees of the generators and relations. In fact, if the genus $g = 0$ or $g = 1$, the proof yields an explicit system of generators and the initial ideal of relations. While in the genus $g \geq 2$ case, the proof does not always yield an explicit system of generators and relations, if the halfcanonical divisor $\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',0,X)$ a log spin curve, and a system of generators and relations for $\halfcan'$ is known, then our proof provides an inductive procedure for determining the generators and initial ideal of relations for $\halfcan$. In fact, such a system of generators and relations are known for $\halfcan'$ in many cases, as detailed in Neves ~\cite{neves:halfcan}[Section III.4].

@@

\begin{prop}
\label{prop:sat_three_induction}
\todo{Make explicit assumptions on genus.}
\todo{Should switch either $R$ and $R'$ or $\sx$ and $\sx'$ to make prime notation consistent (note that inclusion is reversed since this map adds a stacky point to $\sx$; note that we should also say that this map is only ramified at $P$), Aaron: you're right, I wrote the inclusion on rings in the wrong direction. Actually, the map isn't ramified at $P$, it's just defined away from $P$. I decided to remove this from the proposition though.} Let $R_{L'} = R' = k[x]/I'= k[x_3, x_4
, x_5,\ldots, x_m]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $L = L' + \frac{1}{
3}P$, and let $R = R_L$. Suppose $\deg x_i = i$ for $i \in \{3, 4, 5\}$ and that the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_3}(f) < \ord_{x_3}(h) \implies f \prec h.
\end{align*}
Then, the following statements hold.

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ and$ y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_5 \prec 
		y_4 \prec y_3$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x,y_3,y_4,y_5] \\
			&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

@@

\begin{center}
\label{table:g-0-exceptional}
\textbf{Table (IV)}: Genus 0 exceptional cases
\end{center}

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r)$ and has relations in degree at most $\max
(10, e_1, \ldots, e_r),$ so long as $\sigma$ does not lie in a
finite list of exceptional cases, as listed in Table
~\ref{table:g-0-exceptional}.
\end{thm}

\begin{proof}

>>>>

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r).$ If the genus $g \geq 2,$ and $\Delta = 0$ then 
$R(\sx,\Delta, \halfcan)$ has relations in degree at most $\max(11, e_1,
\ldots, e_r).$ Otherwise, $R(\sx, \Delta, \halfcan)$ has relations in degree at most
$\max(10, e_1, \ldots, e_r),$ so long as $\sigma$ does not lie in a finite list of
exceptional cases, as listed in Table ~\ref{table:g-1-exceptional}
for signatures with $g = 1$ and Table ~\ref{table:g_0_exceptional}
for signatures with $g = 0$.
\end{thm}
\todo{Say that this follows immediately from the three theorems for the special cases.}
Note that in the course of the proof of Theorem ~\ref{thm:g_0_generators_relations} we will not merely obtain bounds on the degrees of the generators and relations. In fact, if the genus $g = 0$ or $g = 1$, the proof yields an explicit system of generators and the initial ideal of relations. While in the genus $g \geq 2$ case, the proof does not always yield an explicit system of generators and relations, if the halfcanonical divisor $\halfcan$ can be written as $\halfcan' + \Lambda$ with $(\halfcan',0,X)$ a log spin curve, and a system of generators and relations for $\halfcan'$ is known, then our proof provides an inductive procedure for determining the generators and initial ideal of relations for $\halfcan$. In fact, such a system of generators and relations are known for $\halfcan'$ in many cases, as detailed in Neves ~\cite{neves:halfcan}[Section III.4].

@@

\begin{prop}
\label{prop:sat_three_induction}
\todo{Make explicit assumptions on genus.}
\todo{Should switch either $R$ and $R'$ or $\sx$ and $\sx'$ to make prime notation consistent (note that inclusion is reversed since this map adds a stacky point to $\sx$; note that we should also say that this map is only ramified at $P$), Aaron: you're right, I wrote the inclusion on rings in the wrong direction. Actually, the map isn't ramified at $P$, it's just defined away from $P$. I decided to remove this from the proposition though.} Let $R_{L'} = R' = k[x]/I'= k[x_3, x_4
, x_5,\ldots, x_m]/I'$ satisfy $sat(Eff(\halfcan')) = 3.$ Let $L = L' + \frac{1}{
3}P$, and let $R = R_L$. Suppose $\deg x_i = i$ for $i \in \{3, 4, 5\}$ and that the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_3}(f) < \ord_{x_3}(h) \implies f \prec h.
\end{align*}
Then, the following statements hold.

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ and $y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_5 \prec 
		y_4 \prec y_3$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x,y_3,y_4,y_5] \\
			&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

@@

\begin{center}
\label{table:g_0_exceptional}
\textbf{Table (IV)}: Genus 0 exceptional cases
\end{center}

@@

\noindent
is generated as a $k$-algebra by elements of degree at most $e =
\max(5, e_1, \ldots, e_r)$ and has relations in degree at most $2e$,
so long as $\sigma$ does not lie in a finite list of exceptional
cases, as listed in Table ~\ref{table:g_0_exceptional}.
\end{thm}

\begin{proof}
If $\sx$ has no ramification points, then we have the classical
case which is done in Chapter 4 of Voight and Zureick-Brown
\cite{vzb:stacky}.

Now let us consider the case when $\delta \geq 2$. In this case,
$\halfcan$ is an effective divisor.

Now suppose $\delta = 0$, so $\halfcan$ is not necessarily effective.
If $\sigma = (0; e_1, \ldots, e_r; 0)$ is not one of the
exceptional cases in Table ~\ref{table:g_0_exceptional} and $r \leq
5$, then $\sigma$ ``lies above'' one of the base cases of Lemma
~\ref{lem:g_0_admissible_cases}. In other words, one of the base
cases $(\sigma', J)$ with $\sigma' = (0; e_1', \ldots, e_r'; 0)$
satisfies the property that for all $i \in J$, $e_i \geq e_i'$ and
$e_i = e_i'$ otherwise. If $r > 5$, then we can use Proposition
~\ref{prop:sat_three_induction} to add ramified points with
ramification order $3$ to case (k) of Lemma
~\ref{lem:g_0_admissible_cases}, which corresponds to $\sigma_{11}
= (0; 3, 3, 3, 3, 3; 0), J_{11} = \{1, 2, 3, 4, 5\}$, since the
saturation in this case is equal to $3$.

We know that each base case satisfies all of the hypotheses of
Theorem ~\ref{thm:ram_order_ind} and their corresponding $R' :=
R(\sx', \Delta, \halfcan')$ are generated, as $k$-algebras, up
to degree $e' := \max (5, e_1', \ldots, e_r')$ with relations
generated up to degree $2e'$ by Lemma ~\ref{g_0_admissible_cases}.
It follows from Theorem ~\ref{thm:ram_order_ind}(b) that $R :=
R(\sx, \Delta, \halfcan)$ is generated up to degree $e := \max(5,
e_1, \ldots, e_r)$. It follows from Theorem ~\ref{thm:ram_order_ind}
(d) that the relations are generated up to degree $2e$.
